{\fontfamily{ptm}\selectfont Linear Convergence of Stochastic Nesterov's Accelerated Proximal Gradient method under Interpolation Hypothesis}

### 2025-07-09 Notebook LaTEX/SAGD Under Interpolation/main.tex
>w<, changes

% ==============================================================================
\section{Nesterov's Accelerated Gradient}
    \subsection{In preparations}
        Assumption \ref{ass:smooth-plus-nonsmooth} states the standard assumption of smooth plus non-smooth structured optimization problems for proximal gradient based algorithm. 
        Assumption \ref{ass:smooth-plus-nonsmooth-x} adds the assumption for the existence of minimizers. 
        Definition \ref{def:pg-opt} is the definition of the proximal gradient operator, which is equivalent to the gradient descent operator when the non-smooth part of the objective is the zero function. 
        \par
        Finally, to show the convergence of a stochastic case of the Nesterov's accelerated proximal gradient, we prepared Lemma \ref{thm:jesen} and, \ref{thm:pg-ineq}. 
        They are crucial in the derivation of the convergence. 
        The derivation for the convergence rate of a stochastic accelerated variant of Nester's accelerated  proximal gradient method is in the next section. 
        \begin{assumption}[smooth add nonsmooth]\label{ass:smooth-plus-nonsmooth}
            The function $F = f + g$ where $f:\RR^n \rightarrow \RR$ is an $L$ Lipschitz smooth and $\mu \ge 0$ strongly convex function. 
            The function $g:\RR^n \rightarrow \overline \RR$ is a closed convex proper function. 
        \end{assumption}
        \begin{assumption}[admitting minimizers]\label{ass:smooth-plus-nonsmooth-x}
            Let $F = f + g$ and in addition assume that the set of minimizers $X^+ := \argmin_{x}F(x)$ is non-empty. 
        \end{assumption}

        \begin{definition}[Proximal gradient operator]\label{def:pg-opt}
            Suppose $F = f + g$ satisfies Assumption \ref{ass:smooth-plus-nonsmooth}. 
            Let $\beta > 0$. 
            Then, we define the proximal gradient operator $T_{\beta}$ as 
            \begin{align*}
                T_\beta (x | F) &= \argmin_{z} \left\lbrace
                    g(z) + f(x) + \langle \nabla f(x), z - x\rangle + \frac{\beta}{2}\Vert z - x\Vert^2
                \right\rbrace. 
            \end{align*}
        \end{definition}
        \begin{remark}
            If the function $g \equiv 0$, then it yields the gradient descent operator $T_\beta(x) = x - \beta^{-1}\nabla f(x)$. 
            In the context where it's clear what the function $F = f + g$ is, we simply write $T_\beta(x)$ for short. 
        \end{remark}

        \begin{definition}[Bregman Divergence]
            Let $f:\RR^n \rightarrow \overline \RR$ be a differentiable function. 
            Then, for all the Bregman divergence $D_f: \RR^n \times \dom\nabla f \rightarrow \RR$ is defined as: 
            \begin{align*}
                D_f(x, y) := f(x) - f(y) - \langle \nabla f(y), x - y\rangle. 
            \end{align*}
        \end{definition}
        \begin{remark}
            If, $f$ is $\mu \ge 0$ strongly convex and $L$ Lipschitz smooth then, its Bregman Divergence has for all $x, y \in \RR^n$: $\mu/2 \Vert x - y\Vert^2 \le D_f(x, y) \le L/2 \Vert x - y\Vert^2$. 
        \end{remark}

        % \begin{definition}[R-WAPG sequence]\label{def:rwapg-seq}
        %     Let $(L_k)_{k \ge 0}$ be a sequence such that $L_k > \mu$ for all $k$. 
        %     Let $\alpha_0 \in (0, 1]$, $(\alpha_k)_{k \ge 1}$ has $\alpha_k \in (\mu/ L_k, 1)$. 
        %     Then define for all $k \ge 0$: 
        %     \begin{align*}
        %         \rho_k(1 - \alpha_{k + 1})\alpha_k^2 = \alpha_{k + 1}(\alpha_{k + 1} - \mu/L_k). 
        %     \end{align*}
        % \end{definition}
        % \begin{remark}
        %     When $\rho_k = 1$, the recursive relation between $\alpha_k, \alpha_{k - 1}$ is the same as the well known Nesterov's sequence used in algorithm such as FISTA and Nesterov's accelerated gradient. 
        %     See Li and Wang \cite{li_relaxed_2025} for more information. 
        % \end{remark}
        % \begin{definition}[similar triangle representation of NAPG]\;\label{def:st-method}\\
        %     Let $(\alpha_k)_{k \ge 0}$ be an R-WAPG sequence. 
        %     Suppose that the base case $v_{-1}, x_{k - 1}\in \RR^n$ is given to initialize the algorithm. 
        %     Then the algorithm produces the sequence of iterates $(y_k, x_k, v_k)_{k \ge 0}$ and auxiliary parameter sequence $L_k, \tau_k$ satisfying these inequalities: 
        %     \begin{align*}
        %         & \tau_k = L_k(1 - \alpha_k)(L_k\alpha_k - \mu)^{-1}, 
        %         \\
        %         & y_k = (1 + \tau_k)^{-1} v_{k - 1} + \tau_k (1 + \tau_k)^{-1} x_{k - 1},
        %         \\
        %         & D_f(x_k, y_k) \le L_k/2\Vert x_k - y_k\Vert^2, 
        %         \\
        %         & x_k = T_{L_k}(y_k),
        %         \\
        %         & v_k = x_{k - 1} + \alpha_k^{-1}(x_k - x_{k - 1}). 
        %     \end{align*} 
        % \end{definition}
        % The following theorems are critical in analyzing the behavior of algorithm in Definition \ref{def:st-method}.

@@

        \begin{definition}[semi quasi strong convexity Jensen inequality \textcolor{red}{NEW}]\label{def:sq-scnvx}
            Let $m, n \in \N$. 
            A function $f:\RR^n \rightarrow \overline \RR$ is semi quasi strongly convex with linear mapping $A:\RR^m \rightarrow \RR^n$ if for all $x \in \RR^n, y \in \RR^n$ and $\lambda \in [0, 1]$ it satisfies:

@@

            Choose $A = \sqrt{\mu} I$, where $I$ is the identity matrix then it is equivalent to $\mu$ strongly convex. 
            This is an example of uniform convexity because there exists some $\sigma \ge 0$ such that it fits in Bauschke \cite[Definition 10.7]{bauschke_convex_2017} with $\phi(z) = \sigma^2z^2$. 
            It is called ``Semi'' because of the semi norm induced by matrix $A$, followed by ``Quasi strongly convex'' because the quasi norm $\Vert Ax\Vert^2$ is the distance of an oblique projection to the kernel of $A$ which is a quasi strongly convex function. 
        The above definition is the special case of relatively strong convexity wrt to the Bregman divergence $D_h(x, y) = \Vert Ax - Ay\Vert^2$. 
        \begin{theorem}[an equivalent definition for SQ-SCNVX functions]\label{thm:sq-scnvx-equiv}
            A function $F$ is semi quasi strongly convex with $A$ if and only if the function $x \mapsto F(x) - \frac{1}{2}\Vert A x\Vert^2$ is convex. 

@@

            So, the function $F(x) - \frac{1}{2}\Vert Ax\Vert^2$ is convex. 

@@

@@

        \begin{theorem}[Affine composite differentiable SQ-SCNVX functions]\label{thm:smooth-aff-sq-scnvs-fxn}

        \end{theorem}
        \begin{proof}
            Let $h(x) = f(Ax - b)$.

        \end{proof}
        The following definition defines the concept of relative smoothness. 
        \begin{definition}[semi relative smoothness]\label{def:semi-rel-smooth}

        \end{definition}
        We build the proximal gradient inequality for the class of SQ-SCNVX functions. 
        \begin{theorem}[proximal gradient inequality]\label{thm:pg-ineq}
            Let function $F$ satisfies Assumption \ref{ass:smooth-plus-nonsmooth}, so it's $\mu \ge 0$ strongly convex. 
            For all $x\in \RR^n$, define $x^+ = T_L(x)$, then there exists a $B \ge 0$ such that $D_f(x^+, x) \le B/2 \Vert x^+ - x\Vert^2$. 
            Then, for all $z \in \RR^n$ it satisfies proximal gradient inequality at point $x$:  
            \begin{align*}
                0&\le F(z) - F(x^+) - \frac{B}{2}\Vert z - x^+\Vert^2  + \frac{B - \mu}{2}\Vert z - x\Vert^2
                \\
                &=  F(z) - F(x^+) - \langle B(x - x^+), z - x\rangle
                - \frac{\mu}{2}\Vert z - x\Vert^2
                - \frac{B}{2}\Vert x - x^+\Vert^2. 
            \end{align*}
            Since $f$ is assumed to be $L$ Lipschitz smooth, the above condition is true for all $x, y \in \RR^n$ for all $B \ge L$. 
        \end{theorem}
        \begin{remark}
            The theorem is the same as in Nesterov's book \cite[Theorem 2.2.13]{nesterov_lectures_2018}, but with the use of proximal gradient mapping and proximal gradient instead of project gradient hence making it equivalent to the theorem in Beck's book \cite[Theorem 10.16]{beck_first-order_2017}. 
            The only generalization here is parameter $B$ which made to accommodate algorithm that implements Definition \ref{def:snapg-v2} with line search routine to determine $L_k$. 
            Each of the reference books gives a proof of the theorem. 
            But for the best consistency in notations, see Theorem 2.3 in Li and Wang \cite{li_relaxed_2025}. 
        \end{remark}

@@

    \subsection{convergence rate of the algorithm under various circumstances}

>>>>

% ==============================================================================
\section{Nesterov's Accelerated Gradient}
    \subsection{In preparations}
        Definition \ref{def:pg-opt} is the definition of the proximal gradient operator, which is equivalent to the gradient descent operator when the non-smooth part of the objective is the zero function. 
        To show the convergence of a stochastic case of the Nesterov's accelerated proximal gradient, we prepared Lemma \ref{thm:jesen} and, \ref{thm:pg-ineq}. 
        They are crucial in the derivation of the convergence. 
        The derivation for the convergence rate of a stochastic accelerated variant of Nesterov's accelerated  proximal gradient method is in the next section. 
        \subsubsection{Basic definitions}
            \begin{definition}[Proximal gradient operator]\label{def:pg-opt}
                Suppose $F = f + g$ with $\reli(\dom f) \cap \reli(\dom g) \neq \emptyset$, and $f$ is a differentiable function. 
                Let $\beta > 0$. 
                Then, we define the proximal gradient operator $T_{\beta}$ as 
                    T_\beta (x | F) &= \argmin_{z} \left\lbrace
                        g(z) + f(x) + \langle \nabla f(x), z - x\rangle + \frac{\beta}{2}\Vert z - x\Vert^2
                    \right\rbrace. 
                If the function $g \equiv 0$, then it yields the gradient descent operator $T_\beta(x) = x - \beta^{-1}\nabla f(x)$. 
                In the context where it's clear what the function $F = f + g$ is, we simply write $T_\beta(x)$ for short. 
            \begin{definition}[Bregman Divergence]
                Let $f:\RR^n \rightarrow \overline \RR$ be a differentiable function. 
                Then, for all the Bregman divergence $D_f: \RR^n \times \dom\nabla f \rightarrow \RR$ is defined as: 
                    D_f(x, y) := f(x) - f(y) - \langle \nabla f(y), x - y\rangle. 
                If, $f$ is $\mu \ge 0$ strongly convex and $L$ Lipschitz smooth then, its Bregman Divergence has for all $x, y \in \RR^n$: $\mu/2 \Vert x - y\Vert^2 \le D_f(x, y) \le L/2 \Vert x - y\Vert^2$. 
        \subsubsection{Properties of function}
            \begin{definition}[semi quasi strongly convex function \textcolor{red}{NEW}]\label{def:sq-scnvx}
                A function $F: \RR^n \rightarrow \overline \RR$ is a semi quasi strongly convex function, abbreviated as ``SQ-SCNVX'' with respective to a linear mapping $A \in \RR^{m \times n}$ if $F - \frac{1}{2}\Vert Ax\Vert^2$ is a convex function. 
                Any $\mu \ge 0$ strongly convex function is QS-SCNVX with $A = \sqrt{\mu}I$. 
                But the converse is not true because a seminorm is not a norm. 
                One feature of a QS-SCNVX function is that it doesn't have a unique minimizer which differs it from strong convexity. 
                It may not have a unique minimizer because it's not necessary that $\kernel A = \{\mathbf 0\}$. 

@@

            \begin{theorem}[quasi jesen inequality \textcolor{red}{NEW}]\label{thm:sq-scnvx-equiv}
                A function $F$ is semi quasi strongly convex with $A \in \RR^{m \times n}$ (Definition \ref{def:sq-scnvx}) if and only if, for all $x, y \in \RR^n$ and, $\lambda \in [0, 1]$ it satisfies the inequality: 

@@

@@

                The last line shows that the function $F(x) - \frac{1}{2}\Vert Ax\Vert^2$ is convex, the chain of equality shows the equivalence. 

@@

            \begin{definition}[smoothness and strong convexity with seminorm \textcolor{red}{NEW}]\label{def:seminorm-smooth-scnvx}
                Let $f:\RR^n \rightarrow \RR$ be a differentiable function on $\reli\dom f$. 
                Let $m \in \N$. 
                Let $A: \RR^{m\times n}$ and $b \in \RR^n$ be arbitrary. 
                Then, function $h: \RR^m \rightarrow \RR :=  x\mapsto f(Ax - b)$ is relatively smooth and, relatively strongly convex with respect to the function $x \mapsto (1/2)\Vert A x\Vert^2$: 
                    (\forall x \in \RR^m)(\forall y \in \RR^m)\;
                    \frac{\mu}{2}\Vert Ax - Ay\Vert^2 \le 
                    D_f(x, y) \le \frac{L}{2}\Vert Ax - Ay\Vert^2. 
                The definition exchanged the $\Vert \cdot\Vert^2$ for a seminorm squared: $x \mapsto \Vert Ax\Vert^2$ with some $A: \RR^m \rightarrow \RR^n$. 

@@

            \begin{theorem}[affine composition with strong convexity and smoothness]\label{thm:smooth-aff-sq-scnvs-fxn}
                Let $f: \RR^n \rightarrow \RR$ be an $L$ smooth and, $\mu \ge 0$ strongly convex. 
                Let $A: \RR^m \rightarrow \RR^n$, $b \in \RR^n$ be arbitrary. 
                Let $h: \RR^m \rightarrow \RR= x \mapsto f(Ax - b)$, then the function $h$ satisfies: 
                    (\forall x \in \RR^m)(\forall y \in \RR^m)\; \frac{\mu}{2}\Vert Ax - Ay\Vert^2 
                    \le D_h(x, y) \le \frac{L}{2}\Vert Ax - Ay\Vert^2. 
                Then the Bregman divergence of $h$ is: 
                    D_h(x, y) &= h(x) - h(y) - \langle \nabla h(y), x - y\rangle
                    &= f(Ax - b) - f(Ay - b) - \langle A^T\nabla f(Ay - b), x - y\rangle
                    &= f(Ax - b) - f(Ay - b) - \langle \nabla f(Ay - b), Ax - Ay\rangle
                    &= f(Ax - b) - f(Ay - b) - \langle \nabla f(Ay - b), Ax - b - (Ay - b)\rangle
                    &= D_f(Ax - b, Ay - b). 
                Since $f$ is $L$ smooth and $\mu \ge 0$ strongly convex, it means 
                    \frac{\mu}{2}\Vert Ax - Ay\Vert^2
                    \frac{\mu}{2}\Vert Ax - b - (Ay - b)\Vert^2 
                    &\le D_f(Ax - b, Ay - b)
                    &= D_h(x, y) 
                    &\le \frac{L}{2} \Vert Ax - Ay\Vert^2. 
            The following definition defines the concept of relative smoothness. 
            We build the proximal gradient inequality for the class of SQ-SCNVX functions. 
        \subsubsection{Important inequalities}
            \begin{assumption}[smooth add nonsmooth]\label{ass:smooth-plus-nonsmooth}
                The function $F = f + g$ where $f:\RR^n \rightarrow \RR$ is an $L$ Lipschitz smooth and $\mu \ge 0$ strongly convex function. 
                The function $g:\RR^n \rightarrow \overline \RR$ is a closed convex proper function. 
            \begin{assumption}[admitting minimizers]\label{ass:smooth-plus-nonsmooth-x}
                Let $F = f + g$ and in addition assume that the set of minimizers $X^+ := \argmin_{x}F(x)$ is non-empty. 
            \begin{assumption}[seminorm smooth plus non-smooth]\label{ass:snorm-smth-p-nsmth}
                Let $F = f + g$. Assume that: 
                \begin{enumerate}[noitemsep]
                    \item $f: \RR^m \rightarrow \overline \RR$ is differentiable and, it satisfies Definition \ref{def:seminorm-smooth-scnvx} with $L > \mu \ge 0$ and $A \in \RR^{m \times n}$. 
                    \item $g:\RR^m \mapsto \overline \RR$ is a convex, proper and closed function. 
            \begin{theorem}[proximal gradient inequality]\label{thm:pg-ineq}
                Let function $F$ satisfies Assumption \ref{ass:smooth-plus-nonsmooth}, so it's $\mu \ge 0$ strongly convex. 
                For any $x\in \RR^n$, define $x^+ = T_L(x)$. 
                Then, there exists a $B \ge 0$ such that $D_f(x^+, x) \le B/2 \Vert x^+ - x\Vert^2$ and, for all $z \in \RR^n$ it satisfies the inequality: 
                    0&\le F(z) - F(x^+) - \frac{B}{2}\Vert z - x^+\Vert^2  + \frac{B - \mu}{2}\Vert z - x\Vert^2
                    &=  F(z) - F(x^+) - \langle B(x - x^+), z - x\rangle
                    - \frac{\mu}{2}\Vert z - x\Vert^2
                    - \frac{B}{2}\Vert x - x^+\Vert^2. 
                Since $f$ is assumed to be $L$ Lipschitz smooth, the above condition is true for all $x, y \in \RR^n$ for all $B \ge L$. 
                The theorem is the same as in Nesterov's book \cite[Theorem 2.2.13]{nesterov_lectures_2018}, but with the use of proximal gradient mapping and proximal gradient instead of project gradient hence making it equivalent to the theorem in Beck's book \cite[Theorem 10.16]{beck_first-order_2017}. 
                The only generalization here is parameter $B$ which made to accommodate algorithm that implements Definition \ref{def:snapg-v2} with line search routine to determine $L_k$. 
                Each of the reference books gives a proof of the theorem. 
                But for the best consistency in notations, see Theorem 2.3 in Li and Wang \cite{li_relaxed_2025}. 
            \begin{theorem}[proximal gradient inequality with SQ-SCNVX \textcolor{red}{NEW}]\label{thm:pg-ineq-sq-scnvx}
                Suppose that $F:\RR^m \rightarrow \overline \RR := x \mapsto f(x) + g(x)$ satisfies Assumption \ref{ass:snorm-smth-p-nsmth} with $L > \mu \ge 0$ and $A \in \RR^{m\times n}$. 
                For any $x \in \RR^n$, let $x^+ = T_B(x | F)$.
                Then, there exists some $B \ge 0$ such that $D_f(x^+, x) \le \frac{B}{2}\Vert x - x^+\Vert^2$ and, for all $z \in \RR^m, \eta \in \RR$ it satisfies the inequality: 
                    0 &\le F(z) - F(x^+)
                    + \frac{\eta}{2}\Vert z - x\Vert^2 
                    - \frac{\mu}{2}\Vert Az - Ax\Vert^2
                    + \frac{B - \eta}{2}\Vert z - x\Vert^2
                    - \frac{B}{2}\Vert z - x^+\Vert^2 . 
                Firstly, such a $B > 0$ exists, for example $B = L\Vert A\Vert^2$ would be an option because from Definition \ref{def:seminorm-smooth-scnvx}, it for all $x, y$, $D_f(x, y) \le L/2\Vert Ax - Ay \Vert^2 \le L/2\Vert A\Vert^2\Vert x - y\Vert^2$. 
                But it can be much smaller. 
                The function $z \mapsto g(z) + f(x) + \langle \nabla f(x), z - x\rangle + \frac{B}{2}\Vert z - x\Vert^2$ inside the proximal gradient operator has the minimizer $x^+$.  
                This function is also the sum of a convex, proper closed function $g$ and, a simple quadratic and, it's $B > 0$ strongly convex hence, it satisfies the quadratic growth conditions over its minimizer $x^+ = T_B(x|F)$ so, it follows that for all $z \in \RR^m$: 
                    -\frac{B}{2}\Vert z - x^+\Vert^2 
                    + g(z) + f(x) + \langle \nabla f(x), z - x\rangle + \frac{B}{2}\Vert z - x\Vert^2
                        - g(x^+) - f(x) - \langle \nabla f(x), x^+ - x\rangle - \frac{B}{2}\Vert x^+ - x\Vert^2
                    -\frac{B}{2}\Vert z - x^+\Vert^2 
                    + \left(g(z) + f(z) - f(z) + f(x) + \langle \nabla f(x), z - x\rangle + \frac{B}{2}\Vert z - x\Vert^2\right)
                        +\left(- g(x^+) - f(x^+) + f(x^+) - f(x) - \langle \nabla f(x), x^+ - x\rangle - \frac{B}{2}\Vert x^+ - x\Vert^2\right)
                    -\frac{B}{2}\Vert z - x^+\Vert^2 
                    + \left(F(z) - D_f(z, x) + \frac{B}{2}\Vert z - x\Vert^2\right)
                        + \left(- F(x^+) + D_f(x^+, x) - \frac{B}{2}\Vert x^+ - x\Vert^2\right)
                    &\underset{\text{(a)}}{\le }
                    -\frac{B}{2}\Vert z - x^+\Vert^2 
                    + \left(F(z) - D_f(z, x) + \frac{B}{2}\Vert z - x\Vert^2\right)
                    - F(x^+)
                    - \frac{B}{2}\Vert z - x^+\Vert^2 
                    + F(z) - \frac{\mu}{2}\Vert Az - Ax\Vert^2 
                    + \frac{B}{2}\Vert z - x\Vert^2
                    - F(x^+)
                    &= F(z) - F(x^+)
                        \frac{\eta}{2}\Vert z - x\Vert^2 
                        - \frac{\mu}{2}\Vert Az - Ax\Vert^2
                    + \frac{B - \eta}{2}\Vert z - x\Vert^2
                    - \frac{B}{2}\Vert z - x^+\Vert^2 . 
                At (a), we used the fact that line search asserted the condition $D_f(x^+, x) \le \frac{B}{2}\Vert x^+ - x\Vert^2$. 
                At (b), we used the fact that $f$ satisfies Definition \ref{def:seminorm-smooth-scnvx} so, it has for all $x, z \in \RR^m$, $D_f(z, x) \ge \frac{\mu}{2}\Vert z - x\Vert^2$. 
            \begin{theorem}[seminorm smooth plus non-smooth Jensen \textcolor{red}{NEW}]\label{thm:smnrm-jnsn-smth-nsmth}
                Suppose that $F:\RR^m \rightarrow \overline \RR := x \mapsto f(x) + g(x)$ satisfies Assumption \ref{ass:snorm-smth-p-nsmth} with $L > \mu \ge 0$ and $A \in \RR^{m\times n}$. 
                \todoinline{NOT YET FINISHED}

@@

    \subsection{Convergence rate of the algorithm under various circumstances}
